# Testing the flowtrend method

We're going to assume the `flowtrend()` function has been built, and test it
now.

## 1d example {#1d-example}


Generate data.

```{r 1d-example, eval = FALSE}
set.seed(100)
dt       <- gendat_1d(100, rep(100, 100), offset = 4)
dt_model <- gendat_1d(100, rep(100, 100), return_model = TRUE, offset=4)
ylist = dt %>% dt2ylist()
x = dt %>% pull(time) %>% unique()
plot_1d(ylist)
dt_model %>% select(time, cluster, prob) %>% ggplot() + geom_line(aes(x=time, y=prob, group=cluster, col = cluster))
```

Next, we fit the model.

```{r 1d-example2, eval = FALSE}
set.seed(18)
obj <- flowtrend(ylist = ylist,
                 x = x,
                 maxdev = 1,
                 numclust = 3,
                 l = 2,
                 l_prob = 1,
                 lambda = .1,
                 lambda_prob = .1,
                 verbose = TRUE,
                 nrestart = 5)
 
## Also reorder the cluster labels of the truth, to match the fitted model.
ord = obj$mn[,1,] %>% colSums() %>% order(decreasing=TRUE)
lookup <- setNames(c(1:obj$numclust), ord)
dt_model$cluster = lookup[as.numeric(dt_model$cluster)] %>% as.factor()

## Reorder the cluster lables of the fitted model.
obj = reorder_clust(obj)
```

The data and estimated model are shown here. The dashed lines are the true means.


```{r 1d-example-plot-mean, fig.width = 7, fig.height = 5, eval = FALSE}
plot_1d(ylist = ylist, obj = obj, x = x, add_point = FALSE) +
  geom_line(aes(x = time, y = mean, group = cluster),
            data = dt_model,## %>% subset(time %ni% held_out),
            linetype = "dashed", size=2, alpha = .7)
```

The estimated probabilities are shown here.

```{r 1d-example-plot-prob, fig.width = 7, fig.height = 5, eval = FALSE}
plot_prob(obj=obj, x=x) +
  geom_line(aes(x = time, y = prob, group = cluster, color = cluster),
            data = dt_model, linetype = "dashed")  +
  facet_wrap(~cluster)
```



## Testing monotonicity of objective values

The objective value (that is, the penalized log likelihood) should be monotone
decreasing across EM algorithm iterations.

<!-- (Note: because the admm is imperfect and stops early sometimes, the objective is -->
<!-- sometimes non-monotone. See the example from a few sections up!) -->

```{r test-objective, fig.width = 5, fig.height = 3, eval = TRUE}
testthat::test_that("Objective value decreases over EM iterations.",{
  
  for(iseed in 1:5){
    print(iseed)
    
    ## Generate synthetic data
    set.seed(iseed*100)
    dt       <- gendat_1d(100, rep(10, 100))
    dt_model       <- gendat_1d(100, rep(10, 100), return_model = TRUE)
    ylist = dt %>% dt2ylist()
    x = dt %>% pull(time) %>% unique()
    
    ## Fit model
    obj <- flowtrend_once(ylist = ylist,
                     x = x,
                     maxdev = 5,
                     numclust = 3,
                     lambda = 0.02,
                     l = 1,
                     l_prob = 2,
                     lambda_prob = 0.05)

    ## Test objective monotonicity
    niter_end = length(obj$objective)
    testthat::expect_true(all(diff(obj$objective) < 1E-4))

    ## Make a plot
    g = ggplot(tibble(iter=1:niter_end, objective=obj$objectives)) +
      geom_point(aes(x=iter, y=objective)) +
      geom_line(aes(x=iter, y=objective)) +
      ggtitle(paste0("Seed=", iseed*100))
    print(g)
  }
})
```

## 2d example {#2d-example}

Next, we try out `flowtrend` on a synthetic 2d data example.

```{r 2d-example, fig.width = 10, fig.width = 10, eval = TRUE}
set.seed(100) 
TT = 100
dt <- gendat_2d(TT, rep(100, times = TT))
x = 1:TT
set.seed(10)
obj <- flowtrend_once(ylist = dt$ylist,
                      x = x,
                      maxdev = 3,
                      numclust = 3,
                      l = 2,
                      l_prob = 2,
                      lambda = 0.01,
                      lambda_prob = .01,
                      rho_init = 0.01)
g1 = plot_1d(ylist = dt$ylist, obj = obj, idim = 2) + ylim(c(-7, 4))
g2 = plot_1d(ylist = dt$ylist, obj = obj, idim = 1) + ylim(c(-7, 4))
do.call(ggpubr::ggarrange, c(list(g1, g2), ncol=1, nrow=2))
```

The plot above shows 1d projections in the two dimensions of the response data.

And the plot below are some snapshots in time $t$.

```{r plot-2d-in-1d, fig.width = 8, fig.height = 7}
timelist = c(10, 20, 30, 40, 60, 80)
plist = lapply(timelist, function(tt){
  plot_2d(dt$ylist, obj = obj, tt = tt) + coord_fixed() + ylim(-6, 3) + xlim(-6, 3)
})
do.call(ggpubr::ggarrange, c(plist, ncol=3, nrow=2))
```

## Working with "binned" dataset

Recall that "binning" means we will use binned frequency histogram estimates of
the original particle-level dataset; this is useful when there are many
particles. 


```{r binned}
## library(flowmix)
set.seed(10232)
TT = 100
dt <- gendat_2d(TT, rep(100, times = TT))
manual_grid = flowmix::make_grid(dt$ylist, gridsize = 40)
binres = flowmix::bin_many_cytograms(dt$ylist, manual.grid = manual_grid)
set.seed(100)
obj = flowtrend(ylist = binres$ybin_list,
                countslist = binres$counts_list,
                maxdev = 5,
                numclust = 3,
                l = 2,
                l_prob = 2,
                lambda = .01,
                lambda_prob = .005,
                rho_init = .01,
                nrestart = 1)
obj$objectives %>% plot(type='l')

g1 = plot_1d(ylist = dt$ylist, obj = obj, idim = 2) + ylim(c(-7, 4))
g2 = plot_1d(ylist = dt$ylist, obj = obj, idim = 1) + ylim(c(-7, 4))
do.call(ggpubr::ggarrange, c(list(g1, g2), ncol=1, nrow=2))
```


## Unevenly spaced inputs (x)

Let's try the EM algorithm out with unevenly spaced inputs.

```{r uneven-test, fig.width = 6, fig.height = 6}
set.seed(100)
dt       <- gendat_1d(TT=100, rep(100, times=100))
dt_model       <- gendat_1d(TT=100, rep(100, times=100), return_model = TRUE)
ylist_orig = dt %>% dt2ylist()

ind_rm_list = list(seq(from=10, to=100, by=10),
                   ind_rm = 30:50)
for(ind_rm in ind_rm_list){
  x = (1:100)[-ind_rm]
  ylist = ylist_orig[x]

  set.seed(100)
  obj <- flowtrend_once(ylist = ylist, 
                        x = x,
                        maxdev = 100,
                        numclust = 3,
                        l = 2,
                        l_prob = 2,
                        lambda = 0.1,
                        lambda_prob = 0.1, ## 
                        verbose = TRUE,
                        admm_local_adapt = TRUE,
                        rho_init = 0.01)
  obj$objectives %>% plot(type='o', ylab = "EM objectives")
  plot_1d(obj = obj, ylist = ylist, x = x)
}
```



# Documenting the package and building

We finish by running commands that will document, build, and install the
package.  It may also be a good idea to check the package from within this file.

```{r, results='hide'}
litr::document() # <-- use instead of devtools::document() 
```
